Revisiting the Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation using complex Empirical Orthogonal Functions

### 2022-01-24 analysis/paper/shceof.Rmd
Arregla un par de bugs.

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "r", size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  geom_point(aes(color = part), size = 0.25) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_datetime(NULL, date_labels = "%Y", 
                   breaks = seq(as.POSIXct("1950-01-01"), 
                                as.POSIXct("2020-01-01"), by = "10 years"),
                   minor_breaks = seq(as.POSIXct("1950-01-01"), 
                                      as.POSIXct("2020-01-01"), by = "2 years")) +
  scale_y_continuous(NULL, minor_breaks = NULL, sec.axis = dup_axis()) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"),
        strip.placement = "outside", 
        panel.grid.minor.x = element_line(linetype = 2))  +
  tag_facets("rc") +
  panel_background 
```

@@

```{r lable_placement_min}
lable_placement_min <- function(direction = c("vertical", "horizontal"))  {
  direction <- direction[1]
  function(x, y) {
    if (direction == "vertical") {
      return(which.min(y))
    }
    else if (direction == "horizontal") {
      return(which.min(x))
    }
  }
}
```

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour2(data = ~.x[variable == "o3"], 
                aes(z = estimate, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"], 
                    aes(z = estimate*1e9, label = signif(..level.., 2)),
                    size = 2,
                    skip = 1, stroke = 0.15, stroke.colour = "white", 
                    label.placement = lable_placement_min(),
                    breaks = AnchorBreaks(0, exclude = 0)) +

  scale_fill_divergent_discretised("Air temperature (K)", 
                                   # limits = c(-2.25, 2.25),
                                   labels = one_less,
                                   guide = guide_colorsteps_bottom(15, show.limits = TRUE)) + 
  scale_y_level(minor_breaks= NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_linetype(guide = "none") +
  scale_x_longitude(ticks = 60) +
  facet_grid(~term) +
  axis_labs_smol +
  coord_cartesian(clip = "off") +
  tag_facets()
```

>>>>

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "r", size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  geom_point(aes(color = part), size = 0.25) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_date(NULL, date_labels = "%Y",
               breaks = seq(as.Date("1950-01-01"),
                            as.Date("2020-01-01"), by = "10 years"),
               minor_breaks = seq(as.Date("1950-01-01"),
                                  as.Date("2020-01-01"), by = "2 years")) +
  scale_y_continuous(NULL, minor_breaks = NULL, sec.axis = dup_axis()) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"),
        strip.placement = "outside", 
        panel.grid.minor.x = element_line(linetype = 2))  +
  tag_facets("rc") +
  panel_background 
```

@@

```{r lable_placement_min}
label_placement_min <- function(direction = c("vertical", "horizontal"))  {
  direction <- direction[1]
  function(x, y) {
    browser()
    y <- x[[1]]$y
    x <- x[[1]]$x

    if (direction == "vertical") {
      return(which.min(y))
    }
    else if (direction == "horizontal") {
      return(which.min(x))
    }
  }
}
```

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour2(data = ~.x[variable == "o3"],
                aes(z = estimate, linetype = factor(-sign(..level..))),
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"],
                    aes(z = estimate*1e9, label = signif(..level.., 2)),
                    size = 2,
                    skip = 1, stroke = 0.15, stroke.colour = "white",
                    # label.placement = label_placement_min(),
                    breaks = AnchorBreaks(0, exclude = 0)) +

  scale_fill_divergent_discretised("Air temperature (K)",
                                   # limits = c(-2.25, 2.25),
                                   labels = one_less,
                                   guide = guide_colorsteps_bottom(15, show.limits = TRUE)) +
  scale_y_level(minor_breaks= NULL,
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_linetype(guide = "none") +
  scale_x_longitude(ticks = 60) +
  facet_grid(~term) +
  axis_labs_smol +
  coord_cartesian(clip = "off") +
  tag_facets() +
  NULL
```
